---
title: "Test 3 Extra Credit"
author: "Jake Rozran"
date: "4/19/2021"
output: html_document
---

```{r setup, echo = FALSE, warning = FALSE}
library(knitr)
beer <- c(32, 26, 31, 23, 35, 28, 33, 33, 35, 18, 37, 31, 31, 27, 35)
water <- c(22, 22, 23, 18, 19, 15, 20, 21, 19, 26, 18, 15, 18, 20, 29)

data <- data.frame(beer = beer,
                   water = water)

kable(data)
```

# Assignment:

## State the hypothesis - we want to know if drinking beer causes more mosquitoes to be attracted to you. 

$H_{0}: \mu_{beer} - \mu_{water} = 0$  
$H_{A}: \mu_{beer} - \mu_{water} > 0$

## Draw a histogram of both samples - does this data come from a normal population?

```{r hist, echo = FALSE, warning = FALSE}
library(ggplot2)

plot_dat <- data.frame(value = c(beer, water),
                       variable = c(rep("beer", 15), rep("water", 15)))

ggplot(plot_dat, aes(x = value, fill = variable)) + 
    geom_histogram(position = "identity", alpha = 0.5, bins = 5) + 
    facet_wrap(~ variable) + 
    theme(panel.background = element_blank(), 
          panel.grid.major.y = element_line(color = "light gray"),
          panel.grid.major.x = element_blank(),
          axis.ticks.x = element_blank(),
          axis.ticks.y = element_line(color = "light gray"),
          text = element_text(size = 20), 
          legend.position = "none",
          strip.background = element_blank(), 
          legend.title = element_blank(), 
          axis.title = element_blank())
    
```

## What is the sample mean for both groups?

```{r means, echo = FALSE, warning = FALSE}
beer_mean <- mean(beer)
water_mean <- mean(water)
```

The sample mean for beer is: `r round(beer_mean, 2)`.  
The sample mean for water is: `r round(water_mean, 2)`.  

## How many degrees of freedom are there for each group and total?

```{r df, echo = FALSE, warning = FALSE}
beer_df <- length(beer) - 1
water_df <- length(water) - 1
total_df <- beer_df + water_df
```

Beer degrees of freedom: `r beer_df`  
Water degrees of freedom: `r water_df`  
Total degrees of freedom: `r total_df`  

## What is the sample standard deviation for both groups?

```{r sd, echo = FALSE, warning = FALSE}
beer_sd <- sd(beer)
water_sd <- sd(water)
```

Beer standard deviation: `r round(beer_sd, 2)`  
Water standard deviation: `r round(water_sd, 2)`  

## What is the pooled variance for this experiment?

```{r sp2, echo = FALSE, warning = FALSE}
sp2 <- ((beer_df * beer_sd^2) + (water_df * water_sd^2)) / (beer_df + water_df)
```

$s_{p}^{2} = \frac{df_{beer}s_{beer}^{2} + df_{water}s_{water}^{2}} {df_{beer} + df_{water}}$

The pooled variance is: `r round(sp2, 2)`  

## What is the standard error for this experiment?

```{r se, echo = FALSE}
se = sqrt(sp2 * ((1 / length(beer)) + (1 / length(water))))
```

$SE_{\overline{Y}_{beer}-\overline{Y}_{water}} = \sqrt{s_{p}^{2}(\frac{1}{n_{beer}}+\frac{1}{n_{water}})}$

The standard error is: `r round(se, 2)`

## What is the t-statistic for this experiment?

```{r tstat, echo = FALSE}
tstat <- (beer_mean - water_mean) / se
```

$t = \frac{\overline{Y}_{beer} - \overline{Y}_{wter}} {SE_{\overline{Y}_{beer}-\overline{Y}_{wter}}}$

The t-statistic is: `r round(tstat, 2)`

## What is the critical value of t for this experiment?

```{r tcrit, echo = FALSE}
tcrit <- qt(0.05, total_df, lower.tail = FALSE)
```

The critical value for this experiment is: `r round(tcrit, 2)`

## What is our conclusion? 

Our t-statistic is much more extreme than our critical value, thus we can reject 
the null hypothesis; we have significant evidence that the beer drinking group 
attracts more mosquitos than those who drink water.